---
title: "Time Series Analysis"
author: "Kevin Smith"
date: "November 7, 2014"
output: html_document
---

<head>
<script type="text/javascript" src="js/jquery.min.js"></script>
<script type="text/javascript" src="js/jquery-ui.min.js"></script>
<script type="text/javascript" src="lib/fancyBox/source/jquery.fancybox.pack.js"></script>
<script type="text/javascript" src="js/jquery.tocify.min.js"></script>


<link type="text/css" rel="stylesheet" href="css/jquery.tocify.css" />
<link rel="stylesheet" href="lib/fancyBox/source/jquery.fancybox.css" type="text/css" media="screen" />


<script type="text/javascript">
  $(function($){
    $("#TOC").tocify();
    var addToAll = true;
    var gallery = false;
    var titlePosition = 'inside';
    $(addToAll ? 'img' : 'img.fancybox').each(function(){
      var $this = $(this);
      var title = $this.attr('title');
      var src = $this.attr('data-big') || $this.attr('src');
      var a = $('<a href="#" class="fancybox"></a>').attr('href', src).attr('title', title);
      $this.wrap(a);
    });
    if (gallery)
      $('a.fancybox').attr('rel', 'fancyboxgallery');
    $('a.fancybox').fancybox({
      titlePosition: titlePosition
    });
  });
  
$.noConflict();
</script>

<style type="text/css">
  a.fancybox img {
    border: none;
    -o-transform: scale(1,1); -ms-transform: scale(1,1); -moz-transform: scale(1,1); -webkit-transform: scale(1,1); transform: scale(1,1); -o-transition: all 0.2s ease-in-out; -ms-transition: all 0.2s ease-in-out; -moz-transition: all 0.2s ease-in-out; -webkit-transition: all 0.2s ease-in-out; transition: all 0.2s ease-in-out;
  } 
a.fancybox:hover img {
  position: relative; z-index: 999; -o-transform: scale(1.03,1.03); -ms-transform: scale(1.03,1.03); -moz-transform: scale(1.03,1.03); -webkit-transform: scale(1.03,1.03); transform: scale(1.03,1.03);
}

#TOC {
  position: fixed;
  left: 0;
  top: 0;
  width: 200px;
  overflow:auto;
}

body {
  max-width: 800px;
  margin: auto;
  margin-left:210px;
  line-height: 20px;
}

p { line-height: 1.5; }

.fancybox-title-inside {
  margin-left: 0px !important;
}
</style>

<div id="TOC"></div>




```{r}
plot(runif(10))
```


# Header 1 Wacky Maack
## Header 2
```{r} plot(runif(10)) ```
## Header 2b
```{r} plot(runif(10)) ```

## Header 2c
```{r} plot(runif(10)) ```

## Header 2d
```{r} plot(runif(10)) ```

### HH
```{r} plot(runif(10)) ```

# Header 3
```{r} plot(runif(10)) ```

# Header 3
```{r} plot(runif(10)) ```


```{r} plot(runif(10)) ```

# Header 3
```{r} plot(runif(10)) ```

# Header 3
```{r} plot(runif(10)) ```


# Header 3
```{r} plot(runif(10)) ```


# Header 3
```{r} plot(runif(10)) ```


# Header 3
## HHH
```{r} plot(runif(10)) ```

```{r} plot(runif(10)) ```

```{r} plot(runif(10)) ```


## HHH
# Header 3

# Header 3

# Header 3

# Header 3

## Header 2
## Header 3

